Can we explain the popularity of Netflix Shows?

### 2023-01-14 submission.Rmd
save work

---
title: Predicting Netflix Scores on Google 
author: Thalya Lim
# https://renkun.me/2022/03/06/my-recommendations-of-vs-code-extensions-for-r/
---

```{r}
x <- 1
y <- 2
```

>>>>

---
title: Can we explain the popularity of Netflix Shows? 
author: Thalya Lim
Logo: <Put picture of Netflix logo>
output: 
    rmarkdown::html_document:
        theme: lumen
        highlight: rstudio
# https://renkun.me/2022/03/06/my-recommendations-of-vs-code-extensions-for-r/
# https://haozhu233.github.io/kableExtra/awesome_table_in_html.html
---

# About
This project is about analysing what makes a show on Netflix more likeable?

Are TV shows more well received than Movies?

Are action movies more popular than rom-coms?

Are shows from Korean more popular than Thai shows?

This project aims to answer such questions via going through the whole process of creating a machine learning model

# Packages
Before we get started. We will load all the relevant Pacakages
```{r load packages, message=FALSE}
library(reticulate)
library(DBI)
library(dplyr)
library(lubridate)
library(tidyr)
library(data.table)
library(quanteda)
library(tm)
library(lsa)
library("quanteda.textmodels")
library(ggplot2)
library(kableExtra)
```

I also use a bit of Python for this project, so I will load some Python packages too

```{r echo=FALSE,}
use_python("C:/venv/Scripts/python.exe")
use_virtualenv("C:/venv")
```

```{python}
import pandas as pd
import requests
import urllib
from requests_html import HTML
from requests_html import HTMLSession
import bs4
```

# Getting Data

I use two main data sources for this project, namely Kaggle and Google Search Results

### Source 1: Kaggle 
This dataset provides us with a list of Netflix Shows along with information such as the year it was added, duration, cast, description etc..
There are altogether 8808 Netflix Shows in this dataset.
```{r}
netflixData <- read.csv(file = 'datasets/dataset_main.csv')
```
```{r echo=FALSE}
# FOR Displaying results
# temp = head(netflixData, 10)
# temp %>%
#   kbl() %>%
#   kable_styling() %>%
#   column_spec(1, width = "10cm") %>%
#   scroll_box(width = "1000px", height = "500px")
```

### Source 2: Google Search Results 
To represent popularity, I will use Google Review Scores, this is also going to be my dependent variable. 
In order to supplement the information from the Kaggle dataset, I extracted other information from Google which will hopefully also serve as independent features.

The image below shows an example of the data that will be collected for each of the 8808 shows in the Kaggle Dataset, namely Review Score, Description, Genre and Language

```{r pressure, echo=FALSE, out.width = '100%'}
knitr::include_graphics("Search Example.JPG")
```

#### Google Scrapping Method
```{python}
def get_source(url):
    try:
        session = HTMLSession()
        response = session.get(url)
        return response
    except requests.exceptions.RequestException as e:
        print(e)

def get_googleInfo(title):
    scrapedData = []

    query = urllib.parse.quote_plus(title)
    response = get_source("https://www.google.co.uk/search?q=" + query)
    soup = bs4.BeautifulSoup(response.text,"html.parser")

    rating_object=soup.find_all( "div", {"class":"a19vA"} )
    info_object=soup.find_all( "div", {"class":"rVusze"} )
    desc_object=soup.find_all( "div", {"class":"PZPZlf hb8SAc"} )

    scrapedData.append('Title: ' +title)

    for rating in rating_object:
        ratingInfo = rating.getText()
        rating = ratingInfo[0:3]
        scrapedData.append('Rating: ' +rating)

    for info in info_object:
        info = info.getText()
        if 'Language' in info:
            scrapedData.append(info)
            break

    for info in info_object:
        info = info.getText()
        if 'Genre' in info:
            scrapedData.append(info)
            break

    for desc in desc_object:
        description = desc.getText()
        description = description[0:11] + str(': ') + description[11:]
        scrapedData.append(description)
        break

    return scrapedData
```
Here is an example of Output, when we insert the hit show "Squid Game" into the method above.

``` {python}
info = get_googleInfo("Squid Game")
print(info)
```

Load Google Search Results dataset.
``` {r}
googleData <- read.csv(file = 'datasets/dataset_googleScraped.csv')
```

### Joining the Kaggle Dataset and Google Search Results
I combine both datasets. 
```{r}
con <- dbConnect(RSQLite::SQLite(), ":memory:")

dbWriteTable(con, "netflixData", netflixData)
dbWriteTable(con, "googleData", googleData)

query <- "SELECT netflixData.show_id
        , netflixData.type
        , netflixData.title
        , netflixData.director
        , netflixData.cast
        , netflixData.country
        , netflixData.date_added
        , netflixData.release_year
        , netflixData.rating
        , netflixData.duration
        , netflixData.listed_in
        , netflixData.description
        , googleData.Likes
        , googleData.Languages
        , googleData.Genres
        , googleData.Descriptions
        FROM netflixData left join googleData 
        ON netflixData.show_id= googleData.show_id"
res <- dbSendQuery(con, query)
df <- dbFetch(res)

dbClearResult(res)
```

# Data Transformation
## Renaming Columns
I rename the columns because some of them had the same names. 
Independent features that were scraped from Google have "google_" at the front of them.
``` {r}
new_names <- c("id" 
        , "type"
        , "title"
        , "director"
        , "cast"
        , "country"
        , "date_added"
        , "release_year"
        , "rating"
        , "duration"
        , "genre"
        , "plot"
        , "google_score"
        , "google_language"
        , "google_genre"
        , "google_plot")
df <- setNames(df, new_names)
```
## Data Preview
Here is what the dataset looks like now, we can see all the features that will be used:
``` {r echo=FALSE}
temp = head(df, 10)
temp %>%
  kbl() %>%
  kable_styling() %>%
  column_spec(1, width = "10cm") %>%
  scroll_box(width = "1000px", height = "500px")
```

## Drop rows
As expected, the Google search results of some Netflix Shows did not give us a Google Review Score.
I drop these rows because there is no point keeping rows with no Dependent Feature.

```{r}
df <- df[grep("%", df$google_score),]
```

# Data Exploration

### Missing Data
```{r}
count_blanks <- function(feature) {
    string1 <- feature
    string2 <- '== ""'
    subset = paste(string1, string2)
    subset <- subset(df, eval(parse(text=subset)))
    count <- count(subset)
  return(count)
}

feature <- new_names # retrieving all the column names

NA_Count <- c(
 toString(count_blanks('id'))
, toString(count_blanks('type'))
, toString(count_blanks('title'))
, toString(count_blanks('director'))
, toString(count_blanks('cast'))
, toString(count_blanks('country'))
, toString(count_blanks('date_added'))
, toString(count_blanks('release_year'))
, toString(count_blanks('rating'))
, toString(count_blanks('duration'))
, toString(count_blanks('genre'))
, toString(count_blanks('plot'))
, toString(count_blanks('google_score'))
, toString(count_blanks('google_language'))
, toString(count_blanks('google_genre'))
, toString(count_blanks('google_plot')))
NA_Count <- data.frame(feature, NA_Count)

print(NA_Count)

```

# Cleaning each feature (sub header) 
Look for inconsistencies, duplicates, outliers, etc. 
Changing type and formats
Recatagorize 
+ id - check completeness
+ type - summary 
+ title - check completeness
+ director - summary
+ cast - summary 
+ country - summary
+ date_added - 
+ release_year
+ rating
+ duration
+ genre
+ plot
+ google_score
+ google_language
+ google_genre
+ google_plot

Handle missing data (sub header)
Missing reviews (sub sub header)
Remove rows with no data on reviews
< code >

Analysing missing datapoints (sub sub header)
< code >
Pretty useless to have these features
But I transform the fields into present/absent
Just in case it has predictive capabilities
< code > 
# See if type needs cleaning

``` {r}
view <- df %>% group_by(type) %>% summarise(count = n()) # Show types and count by each type
print(view)
```

director
``` {r}
view <- df %>% group_by(director) %>% summarise(count = n()) # Show directors and count by each director group
print(view)
```
Some of the rows have more than one director (will make maping tables)

# See if country needs cleaning
``` {r}
view <- df %>% group_by(country) %>% summarise(count = n()) # Show country and count by each country group

```
# Many of the rows have more than one country (will make maping tables)

# See if date_added needs cleaning
``` {r}
view <- df %>% group_by(date_added) %>% summarise(count = n()) # Show date_added and count by each date_added
```
# Change to date format
``` {r}
df$date_added <- mdy(df$date_added)
```
# See if release year needs cleaning
``` {r}
view <- df %>% group_by(release_year) %>% summarise(count = n()) # Show release_year and count by each release_year
```
# See if rating needs cleaning
``` {r}

view <- df %>% group_by(rating) %>% summarise(count = n()) # Show release_year and count by each release_year
df$rating[df$rating == "PG-13" ] <- "PG"
df$rating[df$rating == "TV-PG" ] <- "PG"
df$rating[df$rating == "TV-14" ] <- "PG"
df$rating[df$rating == "TV-G" ] <- "G"
df$rating[df$rating == "TV-MA" ] <- "R"
df$rating[df$rating == "TV-Y7" ] <- "Y"
df$rating[df$rating == "TV-Y7-FV" ] <- "Y"
df$rating[df$rating == "TV-Y" ] <- "Y"
df$rating[df$rating == "NR" ] <- "UR"

```
# See if duration needs cleaning
``` {r}
view <- df %>% group_by(duration) %>% summarise(count = n()) # Show duration and count by each release_year
```
# Duration convert blanks to NA
# Convert duration to numeric

``` {r}
df$duration[df$duration ==""] <- NA
df <- separate(df, duration, c("duration", "duration_units"))
df <- transform(df, duration = as.numeric(duration))
```

# See if genre needs cleaning
``` {r}
view <- df %>% group_by(genre) %>% summarise(count = n()) # Show genre and count by each genre
```

# See if plot needs cleaning
``` {r}
view <- df %>% group_by(plot) %>% summarise(count = n()) # Show plot and count by each plot
```

# See if google_score needs cleaning
``` {r}
view <- df %>% group_by(google_score) %>% summarise(count = n()) # Show google_score and count by each google_score
```

# Remove % and convert to numeric type
``` {r}
df$google_score<-gsub("%","",as.character(df$google_score))
df <- transform(df, google_score = as.numeric(google_score))
```

Transform google_language and google_genre feature into present/absent because it was not scrappable
``` {r}
df$google_language <- ifelse(df$google_language =="", "absent", "present")
view <- df %>% group_by(google_language) %>% summarise(count = n())  # Show categories and count by each category
print(view)
```
``` {r}
df$google_genre <- ifelse(df$google_genre =="", "absent", "present")
view <- df %>% group_by(google_genre) %>% summarise(count = n())
print(view)
```

``` {r}
write.csv(df, "datasets/view_main_cleaned.csv", row.names=FALSE)
df <- read.csv(file = 'datasets/view_main_cleaned.csv')
```

# Mapping Tables 
# Build Country Mapping Table
``` {r}

country_mapping <- df %>%
    mutate(country = strsplit(as.character(country), ",")) %>%
    unnest(country)

country_mapping = subset(country_mapping, select = c(id, country))
country_mapping$country <- trimws(country_mapping$country, which = c("left"))

write.csv(country_mapping, "datasets/mapping_country.csv", row.names=FALSE)
```

# # Build Director Mapping Table
``` {r}

director_mapping <- df %>%
    mutate(director = strsplit(as.character(director), ",")) %>%
    unnest(director)

director_mapping = subset(director_mapping, select = c(id, director))
director_mapping$director <- trimws(director_mapping$director , which = c("left"))
write.csv(director_mapping, "datasets/mapping_director.csv", row.names=FALSE)
```

# # Build Cast Mapping Table
``` {r}
cast_mapping <- df %>%
    mutate(cast = strsplit(as.character(cast), ",")) %>%
    unnest(cast)

cast_mapping = subset(cast_mapping, select = c(id, cast))
cast_mapping$cast <- trimws(cast_mapping$cast, which = c("left"))
write.csv(cast_mapping, "datasets/mapping_cast.csv", row.names=FALSE)
```

# # Build Genre Mapping Table
``` {r}
genre_mapping <- df %>%
    mutate(genre = strsplit(as.character(genre), ",")) %>%
    unnest(genre)

genre_mapping = subset(genre_mapping, select = c(id, genre))
genre_mapping$cast <- trimws(genre_mapping$genre, which = c("left"))
write.csv(genre_mapping, "datasets/mapping_genre.csv", row.names=FALSE)
```

# Trim dataset for better performance
``` {r}
df = subset(df, select = -c(country,director,cast,genre) )
```

#Get top 100 actors and actress from IMDB list

# Feature engineering - categories actors based on how popular they are 
``` {r}
popularActors <- read.csv(file = 'datasets/dataset_popular_actors.csv')

# Create new column and set default value to "top 1000"
popularActors$rank_category <- "top 1000"
setDT(popularActors)

popularActors$rank_category[popularActors$Position < 501 ] <- "top 500"
popularActors$rank_category[popularActors$Position < 201 ] <- "top 200"
popularActors$rank_category[popularActors$Position < 101 ] <- "top 100"
popularActors$rank_category[popularActors$Position < 51 ] <- "top 50"
popularActors$rank_category[popularActors$Position < 21 ] <- "top 20"
popularActors$rank_category[popularActors$Position < 11 ] <- "top 10"

popularActors <- separate(popularActors, Description, c("Points", "Description"))
popularActors <- transform(popularActors, Points = as.numeric(Points))

# View(popularActors)

write.csv(popularActors, "datasets/view_popular_actors.csv", row.names=FALSE)
```

# Feature engineering - categories actors based on how popular they are 
``` {r}
popularActresses <- read.csv(file = 'datasets/dataset_popular_actresses.csv')

# Create new column and set default value to "top 1000"
popularActresses$rank_category <- "top 1000"
setDT(popularActresses)

popularActresses$rank_category[popularActresses$Position < 501 ] <- "top 500"
popularActresses$rank_category[popularActresses$Position < 201 ] <- "top 200"
popularActresses$rank_category[popularActresses$Position < 101 ] <- "top 100"
popularActresses$rank_category[popularActresses$Position < 51 ] <- "top 50"
popularActresses$rank_category[popularActresses$Position < 21 ] <- "top 20"
popularActresses$rank_category[popularActresses$Position < 11 ] <- "top 10"

# Get points from actresses
popularActresses <- separate(popularActresses, Description, c("Points", "Description"))
popularActresses <- transform(popularActresses, Points = as.numeric(Points))

# View(popularActresses)

write.csv(popularActresses, "datasets/view_popular_actresses.csv", row.names=FALSE)
```

# Feature engineering - categories directors based on how popular they are 
``` {r}
popularDirectors <- read.csv(file = 'datasets/dataset_popular_directors.csv')

# Create new column and set default value to "top 1000"
popularDirectors$rank_category <- "top 1000"
setDT(popularDirectors)

popularDirectors$rank_category[popularDirectors$Position < 501 ] <- "top 500"
popularDirectors$rank_category[popularDirectors$Position < 201 ] <- "top 200"
popularDirectors$rank_category[popularDirectors$Position < 101 ] <- "top 100"
popularDirectors$rank_category[popularDirectors$Position < 51 ] <- "top 50"
popularDirectors$rank_category[popularDirectors$Position < 21 ] <- "top 20"
popularDirectors$rank_category[popularDirectors$Position < 11 ] <- "top 10"

#Get points from directors
popularDirectors <- separate(popularDirectors, Description, c("Points", "Description"))
popularDirectors <- transform(popularDirectors, Points = as.numeric(Points))

# View(popularDirectors)

write.csv(popularDirectors, "datasets/view_popular_directors.csv", row.names=FALSE)
```

# Create an ephemeral in-memory RSQLite database
#``` {r}

con <- dbConnect(RSQLite::SQLite(), ":memory:")

main <- read.csv(file = 'datasets/view_main_trimmed.csv')
cast <- read.csv(file = 'datasets/mapping_cast.csv')
actorsPoints <- read.csv(file = 'datasets/view_popular_actors.csv')
actressesPoints <- read.csv(file = 'datasets/view_popular_actresses.csv')
directors <- read.csv(file = 'datasets/mapping_director.csv')
directorsPoints <- read.csv(file = 'datasets/view_popular_directors.csv')

dbWriteTable(con, "main", main)
dbWriteTable(con, "cast", cast)
dbWriteTable(con, "actorsPoints", actorsPoints)
dbWriteTable(con, "actressesPoints", actressesPoints)
dbWriteTable(con, "directors", directors)
dbWriteTable(con, "directorsPoints", directorsPoints)

# Attach the points to the mapping table
query <- "SELECT c.id
                ,c.cast
                ,CASE WHEN a.Points IS NOT NULL THEN a.Points
                WHEN b.Points IS NOT NULL THEN b.Points 
                ELSE 0 END AS Points
                FROM cast c
                LEFT JOIN actorsPoints a ON c.cast = a.Name
                LEFT JOIN actressesPoints b ON c.cast = b.Name"
res <- dbSendQuery(con, query)
castPointsMapping <- dbFetch(res)
dbClearResult(res)

dbWriteTable(con, "castPointsMapping", castPointsMapping)

#Attach the director to the mapping table
query <- "SELECT a.id
                ,a.director
                ,CASE WHEN d.Points IS NOT NULL THEN d.Points
                ELSE 0 END AS Points
                FROM directors a
                LEFT JOIN directorsPoints d ON a.director = d.Name"
res <- dbSendQuery(con, query)
directorPointsMapping <- dbFetch(res)
dbClearResult(res)

dbWriteTable(con, "directorPointsMapping", directorPointsMapping)

query <- "SELECT a.*
        , CASE WHEN c.Points IS NOT NULL THEN max(c.Points)
        ELSE 0 END AS cast_points
        , CASE WHEN d.Points IS NOT NULL THEN max(d.Points)
        ELSE 0 END AS director_points
        FROM main a 
        LEFT JOIN castPointsMapping c ON a.id = c.id 
        LEFT JOIN directorPointsMapping d ON a.id = d.id 
        GROUP BY a.id"

res <- dbSendQuery(con, query)
df <- dbFetch(res)

dbClearResult(res)
#```

# Seperate google_scores into bins
#``` {r}
# df$google_scoreClass <- cut(df$google_score, breaks=c(0,60,70,80,90,100), labels=c("0-60","61-70","71-80","81-90","90-100"))

write.csv(df, "datasets/view_main_features.csv", row.names=FALSE)

dbDisconnect(con)
#```

# Latent Semantic Analysis of Plot and google plot data
``` {r}

```
# load data

#``` {r}
main <- read.csv(file = 'datasets/view_main_features.csv')
main$plot_combined <- paste(main$plot,main$google_plot)
#```
# Convert data to corpus
#``` {r}
corpus <- Corpus(VectorSource(main$plot_combined))
#```

# Remove special characters, case, punctuation and numbers
#``` {r}

removeSpecialChars <- function(x) gsub("[^a-zA-Z0-9 ]","",x)
corpus <- tm_map(corpus, removeSpecialChars)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, content_transformer(tolower))
#```

# Remove stop words
#``` {r}
corpus <- tm_map(corpus, removeWords, stopwords("english"))
#```

# Remove white space
#``` {r}
corpus <- tm_map(corpus, stripWhitespace)
#```

# Stemming
#``` {r}
corpus <- tm_map(corpus, stemDocument)
#```

# Term document matrix
#``` {r}
tdm <- TermDocumentMatrix(corpus)
#```

# View top 5 most frequently occuring term
#``` {r}
tdmMatrix <- as.matrix(tdm)
v <- sort(rowSums(tdmMatrix), decreasing=TRUE)
d <- data.frame(word=names(v),freq=v)
print(head(d,5))
#```

# Weighting 
#``` {r}
tdm2 <- as.textmatrix(as.matrix(tdm))
tdm.weighted <- lw_logtf(tdm2)
print(tdm.weighted)
#```

# Latent Semantic Analysis
#``` {r}
lsa <- lsa(tdm.weighted, dims=10)

Tk <- lsa$tk
Sk <- lsa$sk
Dk <- lsa$dk 
#```

# Varimax Rotation
#``` {r}
Tk.varimax <- varimax(Tk)

Dk.rotated <- Dk %*% Tk.varimax$rotmat
Tk.rotated <- Tk %*% Tk.varimax$rotmat 
#```

# Interreting Term and Document Matrices
#``` {r}

sort.loadings.table <- function(x) {
    factor.names <- colnames(x)
    x <- as.data.frame(x)
    x$max.factor <- apply(abs(x),1,which.max)
    x$max.value <- x[cbind(1:nrow(x),x$max.factor)]
    x <- x[order(x$max.factor,-abs(x$max.value)),]
    x$item <- rownames(x)
    rownames(x) <- NULL
    x
}

threshold.loadings.table <- function (x,q=0.1) {
    threshold <- quantile(abs(x$max.value),probs=1-q)
    x <- subset(x,abs(x$max.value)>threshold)
    x
}

Tk.thresholded <- as.data.table(threshold.loadings.table(sort.loadings.table(Tk.rotated),q=0.02))
write.csv(Tk.thresholded[,.(item, max.factor, max.value)], "datasets/tk.csv", row.names=FALSE)

Dk.thresholded <- as.data.table(threshold.loadings.table(sort.loadings.table(Dk.rotated),q=1))
# print(Dk.thresholded[,.(item,max.factor,max.value)])
write.csv(Dk.thresholded[,.(item,max.factor,max.value)],"datasets/dk.csv", row.names=FALSE)
#```
#``` {r}
biplot <- function(Tk,Dk,xdim,ydim,Tk.scale=1,...){
    print("made it")
    Tk[,type:="Terms"]
    Dk[,type:="Title"]
    biplot.data <- rbindlist(list(Tk,Dk))
    setnames(biplot.data,xdim,"F1")
    setnames(biplot.data,ydim,"F2")
    biplot.data[type=="Terms",F1:=F1*Tk.scale]
    biplot.data[type=="Terms",F2:=F2*Tk.scale]
    ggplot(biplot.data, aes(x=F1, y=F2, label=item, col=type)) + geom_text()
}

plot(biplot(Tk=Tk.thresholded,Dk=Dk.thresholded, xdim="V1", ydim="V2", Tk.scale=2))
#```

# Join LSA data to main dataframe

#``` {r}

con <- dbConnect(RSQLite::SQLite(), ":memory:")

main <- read.csv(file = 'datasets/view_main_features.csv')
lsa <- read.csv(file = 'datasets/lsa_2.csv')

lsa$item <- sub("^", "s", lsa$item )

dbWriteTable(con, "main", main)
dbWriteTable(con, "lsa", lsa)

query <- "SELECT m.*, l.*
                FROM main m
                LEFT JOIN lsa l ON m.id = l.item"
res <- dbSendQuery(con, query)
df <- dbFetch(res)
dbClearResult(res)
setDT(df)
View(df)

write.csv(df, "datasets/view_main_lsa.csv", row.names=FALSE)

dbDisconnect(con)
#```

# Machine learning model Multi Regressions
#``` {r}
main <- read.csv(file = 'datasets/view_main_lsa.csv')

main$date_added <- as.Date(as.character(main$date_added))
main$max.factor<-as.character(main$max.factor)

main = subset(main, select = -c(plot,google_plot,id,title,duration_units,item,max.value))

row.number <- sample(1:nrow(main), 0.5*nrow(main))
train = main[row.number,]
test = main[-row.number,]

print("running regression.....")
model1 = lm(log(google_score)~., data=train)
print(summary(model1))
#```
